% !TeX spellcheck = en_US
\chapter{Future Work}

My future research will be focused mainly on learning and implementing the
Diffusion Monte Carlo Method for bosons with a contact interaction, as my
current work on implementing Variational Monte Carlo Method has been very
successful.

Among other goals, the more important are (without priority order):

\begin{itemize}
  \item Calculate the static structure factor and the momentum distribution of a
        Lieb-Liniger gas through VMC and compare with the results obtained from
        the Lieb-Liniger theory.
  \item
  \item Calculate the static structure factor and the momentum distribution for
        a Bose gas within the multi-rods structure.
\end{itemize}


The schedule from the current semester 2017-1 to the 2018-2 is planned as
follows:

\begin{enumerate}
  \item \textbf{Period 2017-1}. Presentation of the candidature exam.
        Implementation of the software to calculate structure factor and
        momentum distribution with Variational Monte Carlo.

        Towards the end of the semester we expect to have a functional version
        of the Diffusion Monte Carlo algorithm to calculate the energy of the
        ground state for the Lieb-Liniger gas in 1D.

  \item \textbf{Period 2017-2}. We plan to implement the \textbf{DMC} method for
        an interacting many-bosons gas within the 1D multi-rods before extend
        the calculation to the Bose gas within multi-slabs. Again, we will
        calculate al least the ground state energy, the structure factor and the
        momentum distribution.

        Towards the end of the semester 2017-2 we expect to have a computer
        program working efficiently to calculate the ground state energy for the
        1D Lieb-Liniger gas using the Diffusion Monte Carlo algorithm. These
        results will be reported.

  \item \textbf{Period 2018-1}. Extend the analysis to an interacting Bose gas
        within multi-slabs, and report results.

  \item \textbf{Period 2018-2}. Write PhD. Thesis. Present the degree exam.
\end{enumerate}
